---
title: "homework-basic-r-proficiency"
author: "Rob Wells"
date: "2024-08-23"
output: html_document
---

# Homework - Basic R Proficiency

This class exercise is a companion to the following chapter in Computational Text Analysis: [Ch. 11  Loading and Analyzing Data](https://wellsdata.github.io/CompText_Jour/docs/r-load-analyze-visualize.html)


#1: load tidyverse, rio and janitor libraries
```{r}
library(tidyverse)
library(rio)
library(janitor)

```


#2: Import Data using read.csv:

https://github.com/wellsdata/CompText_Jour/raw/main/data/mob_analysis_data.csv
```{r}

library(readr)
library(dbplyr)

github_url <- ("https://github.com/wellsdata/CompText_Jour/raw/main/data/mob_analysis_data.csv")


mob_analysis_data <- read_csv("https://github.com/wellsdata/CompText_Jour/raw/main/data/mob_analysis_data.csv") %>% as.data.frame()

```

#3: Provide dimensions: Gives number rows, then columns

```{r}

nrow(mob_analysis_data)
ncol(mob_analysis_data)

```

#4: Show names of your columns

```{r}

glimpse(mob_analysis_data)
```

#5: Check data types

```{r}

str(mob_analysis_data)


```

#6: Display the first six rows

```{r}
head(mob_analysis_data)


```


#7: Create a table counting the number of entries by year and display the top 5 results

```{r}
library(dplyr)

entries_by_years <- mob_analysis_data %>%
  count(Year) %>%
  arrange(desc(Year)) %>%
  slice_max(n, n=5)

  
  


```

#8: Create a table that counts the different entries in the "keyword" column

```{r}
keyword_column <- mob_analysis_data %>%
  count(keyword, sort = TRUE, name = "Count") %>%
  arrange(desc(Count))

```


#9: Create a simple column ggplot chart that shows the total entries by year

```{r}

library(dplyr)
library(ggplot2)


  mob_analysis_data %>%
  group_by(Year) %>%
count(Year) %>%
  ggplot(aes(x=Year, y=n, fill=n))+
  geom_col(position = "dodge")+
labs(title = "Total entries by year",
       subtitle = "SF PD Service Call Data 2017-2019: Source: SFPD",
       caption = "Graphic by Wells")



```

